---
title: "EDA"
author: "Francois Ban"
date: "10/4/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

Import Data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2016-01-01",
                      date_max = "2016-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>% #adds additional variable that gives me the month I care about
  select(name, id, date, month, everything())
weather_df
```

lubricate package - for date packages

##Start grouping

```{r}
weather_df %>% 
  group_by(name, month) #name variables i want to group by; can get how many distinct groups there are (can see this next to #group (36))
```

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n()) #n() function says the count - how many elements r in this group
```

Counting # of elements in each month (i.e. each group)
```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(number = n()) 

weather_df %>% 
  count(month) #same as above
```


##Grouped Summaries
```{r}
month_df = weather_df %>%
  group_by(month, name) %>%
  summarize(number = n(), #na.rm = TRUE gets rid of all missing variables
            num_na = sum(is.na(tmax)), #what is total # of instances that are missing
            mean_tmax = mean(tmax, na.rm = TRUE), #one number summary of mean of all observations
            mean_prec = mean(prcp, na.rm = TRUE),
            median_prcp = median(prcp),
            sd_tmax = sd(tmax)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point()
```

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
geom_point(alpha = .5) +
geom_line(data = month_df, aes(x = month, y = mean_tmax))
```

Tidy is not always best for humans...

```{r}
weather_df %>%
  group_by(month, name) %>%
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  spread(key = name, value = mean_tmax) #untidy the data so that people can read it better
```

##Grouped mutate

```{r}
weather_df %>%
  group_by(month, name) %>%
  mutate(mean_tmax = mean(tmax, na.rm = TRUE)) #takes mean of each of the groups
```

Mean - gives me one number on other side
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(mean_tmax = mean(tmax, na.rm = TRUE), #average temp in central park and repeat for that block of dataset
         centered_tmax = tmax - mean_tmax) %>%#gives all of central park but centered around zero
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```


##Window Functions - computes some function and puts out the same length of numbers (not just one)
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(tmax_rank = min_rank(desc(tmax))) %>%  #use min_rank not rank they give you smallest ranking value when there r ties
  filter(tmax_rank < 2) #gives you the coldest day
```

Using lags

```{r}
weather_df %>%
  group_by(name, month) %>% #every mutate statement u do respects the grouping from here
  mutate(lag_tmax = lag(tmax)) #if data isnt in order, you have to order first because it will take a vector from previous and move it forward
```

##How much does tempertures change from day to day in each area?
```{r}
weather_df %>%
  group_by(name) %>% 
  mutate(one_day_change = tmax - lag(tmax)) %>% 
  summarize(sd_one_day_change = sd(one_day_change, na.rm = TRUE))
```



